---
title: "STAT 184 Final Project" 
author: "Drew Eddy and Seth Johnson"
output: html_notebook
---
## Research Question




## Front Matter


```{r}
#Libraries used
library(mosaic)
library(tidyverse)
library(DataComputing)
```


## Data Retrieval

Below is the code which reads in the 2 csv files used for this analysis, which are called "dataset_1_raw" and "dataset_2_raw" for now.

The first dataset we used is a csv file of 5000 Movies where each instance is a distict movie with additional info like director, budget, revenuse, etc.

The Second dataset is a similar dataset about Movies with additional columns like runtime and release dates.
```{r}
# retrieving both csv datasets
dataset_1_raw = read_csv("movie_data.csv")
dataset_2_raw = read_csv("stat_2.csv")
```


Below are some lines to inspect both datasets using the head() R function. I used the head functions to get a quick view of what an instance of each dataset looks like

```{r}
dataset_1_raw %>%
  head(5)
```
Since the second dataset (dataset_2_raw) contains several variables with long string values, only useful variable to quick view were selected to inspect
```{r}
dataset_2_raw %>%
  select(original_title, revenue, id, genres)%>%
  head(5)

```



The point of the second dataset is to supplement the first so here we alter the second dataset by selecting only the few variables we need renaming the dataset dataset_2_updated. Also use the head() function to get a quick view of what the new dataset looks like.

```{r}
dataset_2_updated <-
  dataset_2_raw%>%
    select(original_title, revenue, release_date, original_language)

dataset_2_updated%>%
  head(5)
```



```{r}
dataset_1_raw<-
  dataset_1_raw%>%
    select(movie_title, imdb_score, director_name, content_rating, title_year, actor_1_name, actor_2_name, actor_3_name, genres, duration, num_critic_for_reviews)
```
Here we remove variables from our main dataset by selecting all variables we want to remain.



```{r}
movie_data<-
  dataset_1_raw%>%
    left_join(dataset_2_updated, by = c("movie_title"= "original_title"))

movie_data%>%
  head(5)
```
Above we finally join the relevant information from dataset_2_updated to the main dataset (dataset_1_raw) to get a complete dataset which we will use from now on (called movie_data)



## Basic Summary Statistics and Overview


Since the research question asks what factors in to top level imdb scores, below is a view of the top 25 movies titles based of their imbd score

```{r}
movie_data%>%
  select(movie_title, imdb_score)%>%
  arrange(desc(imdb_score))%>%
  head(25)
```
Above we can see that we have a duplicate movie in the database, to remove duplicates we used the distinct() function below (the movie 'Towering Inferno' was also removed because the imdb_score was inaccurate). After these cases are removed, the result top 25 movies is followed
```{r}
movie_data <- movie_data%>%
  distinct(movie_title, .keep_all = TRUE)%>%
  filter(movie_title != 'Towering Inferno')

movie_data%>%
  select(movie_title, imdb_score)%>%
  arrange(desc(imdb_score))%>%
  head(25)
```

## The Influence of the Director

One factor that may result in a movie recieving a high imdb score is the director.  Below is a view of that data that shows which directors (top25) have the highest imbd score on average for their movies and how many movies they have directed

```{r}
movie_data%>%
  group_by(director_name)%>%
  summarise(average_imdb_score = mean(imdb_score), number_of_movies = n())%>%
  arrange(desc(average_imdb_score))%>%
  head(25)
```
It seems from the above view that directors who've only directred one movie do well for imdb_scores, to explore this further, below is a plot showing directors charted by the number of movies and imdb scores



```{r}
director_stat<-
  movie_data%>%
    group_by(director_name)%>%
    summarise(average_imdb_score = mean(imdb_score), number_of_movies = n())%>%
    arrange(desc(average_imdb_score))

```
```{r}
ggplot(data = director_stat, aes(x = average_imdb_score, y = number_of_movies))+
  geom_point(data = director_stat)+ ylim(0, 25)
```

This graph provides a better visualization for stats about Directors. While most of the top imdb scored movies are directed by single movie directors, the above graph shows that there is a positive trend between the number of movies a Director directs and their average_imdb_score, so it might be useful to incorporate the director in the final chart.


## Does the Content Rating influence imdb score


Another possible factor that might correlate with high or low imdb scores is a movie's content rating, which classifies movies based on how mature its content it.  To get a better idea of which content-ratings typically recieve what kind of score, below is a view that shows the average IMDB score by rating. Since we only wanted to include the standard film content rating (G, PG, PG-13, R, X, and Unrated), we used the grepl function and a regular expression to other filter the others out.
```{r}
movie_data<-
  movie_data%>%
  mutate(content_rating = gsub(pattern = "Not Rated", replacement = "Unrated", content_rating))
```

```{r}
movie_data%>%
  select(content_rating, imdb_score)%>%
  group_by(content_rating)%>%
  summarise(average_imdb_score = mean(imdb_score), number_of_movies = n())%>%
  filter(! grepl("TV|NC-17|Approved|Passed|GP|M", content_rating))%>%
  arrange(desc(average_imdb_score))
```
Below is a chart that better shows this correlation better

```{r}
movie_data%>%
  select(content_rating, imdb_score)%>%
  group_by(content_rating)%>%
  summarise(average_imdb_score = mean(imdb_score), number_of_movies = n())%>%
  filter(! grepl("TV|NC-17|Approved|Passed|GP|M", content_rating))%>%
  arrange(desc(average_imdb_score))%>%
  ggplot(aes(x = content_rating, y = average_imdb_score)) +
  geom_histogram(stat = "identity")+
  ylim(0, 7)
```



As we can see the content_rating doesn't seem to have much impact on the imdb score so it probably should have a negligible use in our final chart.


## Multi-Variable Graph

To get a detailed view of which movies recieve higher imdbs, it would be useful to create a chart that incorporates multiple variables and there draw conclusions.


